---
title: "MA615 Tidyverse Problem Set"
author: "Shixin Liang"
date: "10/3/2019"
output: pdf_document
---

```{r setup, include=FALSE}
library(knitr)                ## loading the entire knitr package
library(ggplot2)              ## the only tidyverse library you need is ggplot2
library(esquisse)             ## use esquisse to setup the basic plots
library(tidyr)
library(tidyverse)
library(dplyr)

library(kableExtra)
library(magrittr)
knitr::opts_chunk$set(echo = FALSE)
```
  
  
# Problem 1  
```{r}
library(gapminder)
data(gapminder)
```
  
How many continents are included in the data set?
```{r}
length(levels(gapminder$continent))
```
  
How many countrys are included?
```{r}
length(levels(gapminder$country))
```
  
How many countries per continent?
```{r}
#For Africa
length(unique(gapminder$country[gapminder$continent=="Africa"]))
#For Americas
length(unique(gapminder$country[gapminder$continent=="Americas"]))
#For Asia
length(unique(gapminder$country[gapminder$continent=="Asia"]))
#For Europe
length(unique(gapminder$country[gapminder$continent=="Europe"]))
#For Oceania
length(unique(gapminder$country[gapminder$continent=="Oceania"]))
```
  
A report showing the continents in the dataset, total population per continent, and GDP per capita.
```{r}
gapm <- as_tibble(gapminder)
gapm$pop <-as.numeric(gapm$pop)
# total population per continent
total_pop <- gapm %>% 
  count(continent, wt = pop)
# GDP per capita
total_GDP <- gapm %>% 
  count(continent, wt = gdpPercap)

total <- cbind(total_pop, total_GDP$n)
knitr::kable(total)
```

Produce a well-labeled table that summarizes GDP per capita for the countries in each continent, contrasting the years 1952 and 2007.
```{r}
# calculate mean GDP for each continent
gapm1 <- gapm %>% 
    group_by(year, continent) %>% 
          summarise(mean(gdpPercap))

gapm1_1952 <- filter(gapm1, year == 1952)
gapm1_2007 <- filter(gapm1, year == 2007)
gapm1_1952_2007_table <- rbind(gapm1_1952, gapm1_2007)
knitr::kable(gapm1_1952_2007_table)
```  

Product a plot that summarizes the same data as the table.  
```{r}
ggplot(gapm1_1952_2007_table, aes(year, `mean(gdpPercap)`)) + 
  geom_line(aes(group = continent), colour = "grey50") + 
  geom_point(aes(colour = continent))
```  

Which countries in the dataset have had periods of negative population growth?  
Here are the countries:  
```{r}
pop_change <- gapminder$pop[2:length(gapminder$pop)] - gapminder$pop[1:(length(gapminder$pop)-1)]
pop_change <- c(0, pop_change)
new_gapm <- gapminder
new_gapm$popular_change <- pop_change
negative_grouth <- filter(new_gapm, new_gapm$popular_change <0)
knitr::kable(negative_grouth)

negative_grouth_countries <- unique(negative_grouth$country)
negative_grouth_countries
```  

Which countries in the dataset have had the highest rate of growth in per capita GDP?
```{r}
GDP_change <- gapminder$gdpPercap[2:length(gapminder$pop)] - gapminder$gdpPercap[1:(length(gapminder$pop)-1)]
GDP_change <- c(0, GDP_change)
new_gapm$GDP_change <- GDP_change
highest_rate <- filter(new_gapm, new_gapm$GDP_change == max(new_gapm$GDP_change))
knitr::kable(highest_rate)
```
Kuwait had the highest rate.  

  
  
# Problem 2  
Product a plot the contracts the frequency of these four gendercombinations.
```{r}
library(AER)
data("Fertility")
fert <- unite(Fertility, g1g2, gender1, gender2)
unique(as.factor(fert$g1g2))

# Compute the frequency
freq <- fert %>%
  group_by(g1g2) %>%
  summarise(counts = n())

# Create the bar plot
ggplot(freq, aes(x = g1g2, y = counts)) +
  geom_bar(fill = "#0073C2FF", stat = "identity") +
  geom_text(aes(label = counts), vjust = -0.3) 
```  

Are the frequencies different for women in their 20s and wemen who are older than 29?  
```{r}
fert_20 <- filter(fert, fert$age<30)
fert_30 <- filter(fert, fert$age>=30)

freq_20 <- fert_20 %>%
  group_by(g1g2) %>%
  summarise(counts = n())
freq_20$age <- "20s"
freq_30 <- fert_30 %>%
  group_by(g1g2) %>%
  summarise(counts = n())
freq_30$age <- "30s"
freq_2030 <- rbind(freq_20,freq_30)
knitr::kable(freq_2030)
```
The frequencies for women who are older than 29 are higher.  

  
Produce a plot that contrasts the frequency of having more than two children by race and ethnicity.  
```{r}
fert_race <- unite(Fertility, race, afam, hispanic, other)
fert_race$race[fert_race$race == "no_no_no"] <- "other"
fert_race$race[fert_race$race == "no_no_yes"] <- "Caucasian"
fert_race$race[fert_race$race == "yes_no_no"] <- "African-American"
fert_race$race[fert_race$race == "no_yes_no"] <- "Hispanic"
fert_race$race[fert_race$race == "yes_yes_no"] <- "African-American-Hispanic"
fert_race$race[fert_race$race == "no_yes_yes"] <- "Hispanic-Caucasian"
unique(as.factor(fert_race$race))

fert_race$morekids <- as.character(fert_race$morekids)
fert_race$morekids[fert_race$morekids == "no"] <- 0
fert_race$morekids[fert_race$morekids == "yes"] <- 1
freq_race <- fert_race %>%
  group_by(race) %>%
  summarise(counts = n())

ggplot(freq_race, aes(x = race, y = counts)) +
  geom_bar(fill = "#0073C2FF", stat = "identity") +
  geom_text(aes(label = counts), vjust = -0.3) 

```
  
  
# Problem 3
```{r}
library(stringr)
data(mtcars)
data(mpg)
```  

How many times does the letter “e” occur in mtcars rownames?  
```{r}
mtc_rownames <- rownames(mtcars)
pattern <- "e"
grep(pattern, mtc_rownames, value = TRUE)
occur_time <- sum(str_count(mtc_rownames, pattern))
occur_time
```  

How many cars in mtcars have the brand Merc?  
```{r}
Merc_number <- sum(str_count(mtc_rownames, "Merc"))
Merc_number
```  

How many cars in mpg have the brand(“manufacturer” in mpg) Merc?  
```{r}
mercury_number <- sum(str_count(mpg$manufacturer, "merc"))
mercury_number
```  

Contrast the mileage data for Merc cars as reported in mtcars and mpg. Use tables, plots, and a short explaination.  
```{r}
mtcars_location <-grep("Merc", mtc_rownames)
mile_mtcars <- mtcars$mpg[c(mtcars_location)]

mpg_location <-grep("merc", mpg$manufacturer)
mile_mpg <- mpg$cty[c(mpg_location)]

mile_data <- c(mile_mtcars, mile_mpg)
length(mile_mtcars)
mileage <- data.frame(mile_data)
mileage$data_source <- mileage$mile_data
mileage$data_source[1:length(mile_mtcars)] <- "mtcars"
mileage$data_source[(length(mile_mtcars)+1):(length(mile_mtcars)+length(mile_mpg))] <- "mpg"
mileage$observation <- 1:length(mileage$mile_data)
ggplot(mileage, aes(observation, mile_data)) + 
  geom_point(aes(colour = data_source))

```
The mileage data for Merc cars from  mpg  is smaller than those from mtcars. It is because the mileage data from mpg are mileage data in city, not in highway.  

  
  
# Problem 4
```{r}
library(babynames)
data(babynames)
babyn <- as_tibble(babynames)
```  

Produce a tabble that displays the five most popular boy names and girl names in the years 1880,1920, 1960, 2000.  
```{r}
# boy
# 1880
boy_1880 <- filter(babyn, babyn$year == 1880 & babyn$sex == "M")
boy_name_1880 <- boy_1880$name[1:5]
girl_1880 <- filter(babyn, babyn$year == 1880 & babyn$sex == "F")
girl_name_1880 <- girl_1880$name[1:5]

# 1920
boy_1920 <- filter(babyn, babyn$year == 1920 & babyn$sex == "M")
boy_name_1920 <- boy_1920$name[1:5]
girl_1920 <- filter(babyn, babyn$year == 1920 & babyn$sex == "F")
girl_name_1920 <- girl_1920$name[1:5]

# 1960
boy_1960 <- filter(babyn, babyn$year == 1960 & babyn$sex == "M")
boy_name_1960 <- boy_1960$name[1:5]
girl_1960 <- filter(babyn, babyn$year == 1960 & babyn$sex == "F")
girl_name_1960 <- girl_1960$name[1:5]

# 2000
boy_2000 <- filter(babyn, babyn$year == 2000 & babyn$sex == "M")
boy_name_2000 <- boy_2000$name[1:5]
girl_2000 <- filter(babyn, babyn$year == 2000 & babyn$sex == "F")
girl_name_2000 <- girl_2000$name[1:5]

gender_year <- c("1st", "2nd", "3rd", "4th", "5th")
five_popular_names <- rbind(gender_year, boy_name_1880, girl_name_1880, 
                            boy_name_1920, girl_name_1920, boy_name_1960, girl_name_1960, 
                            boy_name_2000, girl_name_2000)  
knitr::kable(five_popular_names)  
```  

What names overlap boys and girls?  
```{r}
boy <- subset(babynames, babynames$sex=="M", select = name)
girl <- subset(babynames, babynames$sex=="F", select = name)
overlap <- inner_join(boy, girl, by = "name")
overlap <- unique(overlap)
overlap
```  
  
What names were used in the 19th century but have not been used in the 21sth century?  
```{r}
names_19 <- filter(babynames,babynames$year<1900)
names_19 <- unique(names_19$name)
names_21 <- filter(babynames,babynames$year>=2000)
names_21 <- unique(names_21$name)
name_list <- c()
for (i in names_19){
  if (sum(grep(i, names_21))==0){
    name_list <- c(name_list,i)
    }
} 
unique(name_list)  
```

Produce a chart that shows the relative frequency of the names “Donald”, “Hilary”, “Hillary”, “Joe”, “Barrack”, over the years 1880 through 2017.  
```{r}
five_name_freq <- filter(babynames, name == "Donald" | name == "Hilary" | name == "Hillary" | name == "Joe" | name == "Barrack" )
five_name_data <- five_name_freq %>%
  group_by(year, name) %>%
  summarise(sum(n))

ggplot(five_name_data, aes(year, `sum(n)`)) + 
  geom_line(aes(group = name), colour = "grey50") + 
  geom_point(aes(colour = name))
```